### 2011-03-20 extra/sliver_pldi11/abstract.tex

Many static analyses do not scale as they are made more precise.  For example,
increasing the amount of context sensitivity in a $k$-limited analysis causes
the number of contexts to grow exponentially.  Iterative refinement techniques
can be employed to mitigate this growth by starting with a coarse abstraction and
only refining parts of the abstraction deemed relevant with respect to a client.

In this paper, we introduce a new way to use this client feedback to greatly
reduce the complexity of an analysis: {\em pruning}.  The pruned analysis is no
longer a valid abstraction, but we prove that we do not forfeit correctness and
that our analysis is still sound with respect to a given client.  In the
context of $k$-limited analysis, our approach amounts to keeping track of a set
of context prefix patterns.  By iteratively refining and pruning these
patterns, we show that our analysis increases scalability greatly.

>>>>

Many static analyses do not scale as they are made more precise.  For example,
increasing the amount of context sensitivity in a $k$-limited pointer analysis causes
the number of contexts to grow exponentially with $k$.  Iterative refinement techniques
can mitigate this growth by starting with a coarse abstraction and
only refining parts of the abstraction that are deemed relevant with respect to a given client.

In this paper, we introduce a new technique called {\em pruning} that uses client
feedback in a different way.  The basic idea is to use coarse
abstractions to prune away parts of the program analysis deemed irrelevant for proving a
client query, and then using finer abstractions on the sliced program analysis.
%In contrast to existing iterative refinement techniques,
%our analyses do not correspond to abstractions on the original program but rather one on a sliced program.
%We prove that we do not forfeit correctness and
%that our analysis is still sound with respect to a given client. 
For a $k$-limited pointer analysis, this approach amounts to adaptively refining and pruning a set
of prefix patterns representing the contexts relevant for the client.
By pruning, we are able to scale up to much more expensive abstractions than before.
We also prove that the pruned analysis is both sound and complete, that is, it yields the same results
as an analysis that uses a more expensive abstraction directly without pruning.

### 2011-03-20 extra/sliver_pldi11/theory.tex

Recall that the central operation of a static analysis is $\bP$, which serves two roles: (i)
determining if the query is proven (when $\bP$ returns $\emptyset$); and (ii)
returning the subset of relevant input tuples.
The following theorem provides the key equation that drives everything in this paper (see \refapp{proofs} for the proof):

@@

% Generalization
To obtain some intuition behind pruning,
consider the following simpler idea:
first run the analysis with $\beta$; if the query is proven, stop and declare {\em proven}; otherwise, we run the analysis
with $\alpha$ and output that answer.
It is easy to see that this two-step procedure returns the same answer as just running $\alpha$,
because $\beta \preceq \alpha$ implies that if $\beta$ proves the query, then so does $\alpha$
(\refprop{soundness}).
\refeqn{master} can be thought of as an extension of this basic idea:
instead of using $\beta$ to just determine where the query tuple is proven,
we obtain more information, namely the set of input tuples that are relevant.

@@

% Basic
This algorithm essentially applies \refeqn{master} repeatedly.
We first present a simplified version of the algorithm which
ignores the pre-pruning step (we take $A_t = A_t'$).
We are given a sequence of successively finer abstractions $\alpha_0, \alpha_1, \dots$
(e.g., $\alpha_t = \pi_t$ for the $k$-limited abstractions)
and an input set of abstract tuples $A_0$, which is computed
under the initial abstraction $\alpha_0$.  Then the algorithm alternates between a pruning step
and a refining step, always maintaining only the abstract tuples
that could participate in a derivation of the query tuple $\xo$.
On iteration $t$, our current input tuples $A_t$ are first pruned to $\tilde A_t$ using $\bP$;
this is subsequently refined to $A_{t+1}$.
% Example
\reffig{graphIters} shows an example of running this algorithm on the graph
example from \reffig{graphExample}; the first pruning step is shown in
\reffig{graphDerivation}.

@@

% Pre-pruning
Now we discuss pre-pruning.  Pre-pruning requires the user to provide
another sequence of successively finer abstractions $\beta_0, \beta_1, \dots$
which are coarser than $\alpha_0, \alpha_1, \dots$, respectively.
These abstractions will also be used to prune the input tuples.
The idea is that before refining $A_t$ to $A_{t+1}$,
we perform two steps of pruning:
(i) first we use $\beta_t$ in a pre-pruning step;
(ii) then we use $\alpha_t$ during the main pruning step.
Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
the purpose is solely to make the pruning step faster.

Pre-pruning requires a sequence of auxiliary abstractions $(\beta_t)$ which are coarser than the main abstractions $(\alpha_t)$.
We can obtain auxiliary abstractions by composing the main abstractions
with another abstraction $\hclass$ ($\beta_t = \alpha_t \circ \hclass$).
We can use any $\hclass$ for which $\alpha_t \circ \hclass$ yields a valid abstraction,
but the speedup we obtain from pre-pruning depends on the relationship between $\hclass$ and $\alpha_t$.
If $\hclass$ is the total abstraction ($\hclass(c) = \sC$), then pre-pruning will be fast but nothing will be pre-pruned,
so we get no speedup.
If $\hclass$ is no abstraction ($\hclass(c) = c$), then pre-pruning is equivalent to just running the pruning step,
so we again get no speedup.
A good rule of thumb is that $\hclass$ should be ``complementary'' to $\alpha_t$
(we will see some examples in \refsec{abstractions}).

\refthm{algorithm} states that the Prune-Refine algorithm is both
sound and complete.  In other words, pruning has no impact on our answer to a
query---we just compute it more efficiently.
The proof is given in \refapp{proofs} and uses \refthm{master}.

\begin{theorem}[Correctness of the Prune-Refine algorithm]
\label{thm:algorithm}
At iteration $t$,
the incrementally pruned abstraction $A_t$ is equivalent
to the full abstraction $\alpha_t(X)$ (formally, $\bP(\alpha_t(X)) = \bP(A_t)$).
If the algorithm returns ``proven,'' then $\bP(X) = \emptyset$ (the query is actually false).
\end{theorem}

>>>>

Recall that the central operation of a static analysis is $\bP$, which serves two roles: (i)
determining if the query is proven (when $\bP$ returns $\emptyset$); and (ii)
returning the relevant input tuples.
The following theorem provides the key equation that drives everything in this paper (see \refapp{proofs} for the proof):

@@

% Generalization
To obtain some intuition behind pruning,
consider the following simpler idea:
first run the analysis with $\beta$; if the query is proven, stop and declare {\em proven}; otherwise, run the analysis
with $\alpha$ and output that answer.
It is easy to see that this two-step procedure returns the same answer as just running $\alpha$:
Because $\beta \preceq \alpha$, if $\beta$ proves the query, then so does $\alpha$
(\refprop{soundness}).
\refeqn{master} can be thought of as an extension of this basic idea:
instead of using $\beta$ to just determine whether the query tuple is proven,
we obtain more information, namely the whole set of input tuples that are relevant.

@@

% Basic
This algorithm essentially applies \refeqn{master} repeatedly.
We first present a simplified version of the algorithm which
ignores the pre-pruning step (we take $A_t = A_t'$).
We are given a sequence of successively finer abstractions $\alpha_0, \alpha_1, \dots$
(e.g., $\alpha_t = \pi_t$ for the $k$-limited abstractions)
and a set of input abstract tuples $A_0$, which is computed
under the initial abstraction $\alpha_0$.  Then the algorithm alternates between a pruning step
and a refining step, maintaining only the abstract input tuples
that could participate in a derivation of the query tuple $\xo$.
On iteration $t$, our current input tuples $A_t$ are first pruned to $\tilde A_t$ using $\bP$;
this is subsequently refined to $A_{t+1}$.
% Example
\reffig{graphIters} shows an example of running this algorithm on the graph
example from \reffig{graphExample}; the first pruning step is shown in
\reffig{graphDerivation}.

@@

% Pre-pruning
Now we discuss pre-pruning.  Pre-pruning requires the user to provide
another sequence of successively finer abstractions $\beta_0, \beta_1, \dots$
which are coarser than $\alpha_0, \alpha_1, \dots$, respectively.
These abstractions will also be used to prune the input tuples.
The idea is that before refining $A_t$ to $A_{t+1}$,
we perform two steps of pruning:
(i) first we use $\beta_t$ in a pre-pruning step;
(ii) then we use $\alpha_t$ during the main pruning step.
%Pre-pruning does not affect the output of the pruning step ($\tilde A_t$);
%the purpose is solely to make the pruning step faster.

Pre-pruning requires a sequence of auxiliary abstractions $(\beta_t)$ which are coarser than the main abstractions $(\alpha_t)$.
A standard way to obtain auxiliary abstractions is by composing the main abstractions
with another abstraction $\hclass$; formally, $\beta_t = \alpha_t \circ \hclass$.
We can use any $\hclass$ for which $\alpha_t \circ \hclass$ yields a valid abstraction,
but the speedup we obtain from pre-pruning depends on the relationship between $\hclass$ and $\alpha_t$.
If $\hclass$ is the total abstraction ($\hclass(c) = \sC$), then pre-pruning will be fast but nothing will be pre-pruned,
so we get no speedup.
If $\hclass$ is no abstraction ($\hclass(c) = c$), then pre-pruning is equivalent to just running the pruning step,
so we again get no speedup.
A good rule of thumb is that $\hclass$ should be ``complementary'' to $\alpha_t$
(we will see some examples in \refsec{abstractionExamples}).

\refthm{algorithm} states that the Prune-Refine algorithm is both
sound and complete.  In other words, pruning has no impact on the answer to a
query.
The proof is given in \refapp{proofs} and uses \refthm{master}.

\begin{theorem}[Correctness of the Prune-Refine algorithm]
\label{thm:algorithm}
At iteration $t$,
using the incrementally pruned abstraction $A_t$ is equivalent
to using the full abstraction $\alpha_t(X)$ in that $\bP(\alpha_t(X)) = \bP(A_t)$.
Consequently, if the algorithm returns ``proven,'' then $\bP(X) = \emptyset$ (the query is actually false).
\end{theorem}
